R Notebook

### 2017-05-19 manuscript/draft.Rmd
some discussion edits

Forecasting is an emerging imperative in ecology and as such the field needs to
develop and follow good practices for conducting and evaluating ecological
forecasts. We have laid out a number of these practices (Box 1) and attempted to
implement them in a single study that builds and evaluates forecasts of
biodiversity in the form of species richness. The results of this effort are
both promising and humbling. On an absolute basis, when comparing forecasts
across sites, many different approaches to forecasting produce reasonable
forecasts. If a site is predicted to have a high number of species in the
future, relative to other sites, it generally does. However, within a site the
forecasts perform relatively poorly. If a site is predicted to have a higher
richness next year than it did last year there is little confidence that this
will be the case. This results in simple site-level time-series and baseline
models like the long-term average and the naive model being some of the best
performing overall models for future biodiversity.

The most commonly used method to forecast the future state of biodiversity,
fitting a species distribution model separately to each species and then summing
the probability of presence across those models, provides the worst overall
forecasts of all methods evaluated. Most forecasting based on this approach is
never evaluated and when it is it is typically only assessed based on cross-site
rather than within site approaches. As a result the forecasts may appear to
perform well even when change through time is not accurately predicted. This
issue has also recently been recently highlighted as an issue for using SDMs to
predict the future locations of individual species [plos one 2012 paper]. This
result is particularly sobering because this method forms the foundation for
major claims regarding the predicted loss of biodiversity to climate change
[2004 science/nature papers]. The poor performance of SDM-based forecasts
highlights the crucial importance of comparing multiple modeling approaches when
conducting ecological forecasts, and in particular the value of comparing
results to simple baselines to avoid over-interpreting the accuracy of
ecological forecasts. Disciplines that have more mature forecasting cultures
often do this by reporting "forecast skill", which is the improvement in the
forecast relative to a simple baseline. We recommend adopting this approach in
future ecological forecasting research. It also highlights the need for using
hindcasting and time-series data for assessing the effectiveness of forecasting
methods, since this allows comparisons to baselines and assessment of the
predicted dynamics at each site.

The most common alternative to summed SDMs is modeling richness directly using
environmental variables and using the resulting models to make forecasts [].
Two comparisons of the the SDM and direct richness modeling approaches reported
that the methods yielded equivalent results for forecasting diversity. While our
results generally support this for point estimates of richness they also show
that modeling richness directly results in much better estimation of the
uncertainty of the forecast and therefore is a better overall approach to
forecasting richness. A similar result is seen when comparing joint species
distribution models (JSDMs) to stacked single species distribution models. For
point estimates the joint distribution models are roughly equivalent to stacked
SDMs, but the JSDMs provide much better estimates of uncertainty. In fact, JSDMs
and modeling richness directly provide some of the best estimates of uncertainty
across all modeling approaches. This highlights the importance evaluating models
based on their uncertainty as well as their mean estimates. Not doing so is
particularly problematic in this case because the summed species distribution
models are overly confident meaning that using them would lead to believing that
richness would be restricted to a much narrower range than would actually
occur.

The accurate estimates of uncertainty by JSDMs and richness models result in
these models approaching the performance of time-series models and baselines at
the longest time-scales assessed when using measures that consider both point
estimates and uncertainty. This convergence at longer time scales highlights the
importance of considering how forecasting method performance changes with the
distance into the future being predicted. The value of this approach, and
associated forecast horizons, has been raised for knowing how far into the
future a model can be effectively used for forecasting [@petchy2015], but our
results suggest a broader value to this approach for considering the potential
importance of different models and processes for making forecasts at different
scales. [[Should make sure Petchy et al. don't make this point]] Our results
show that for forecasts 1-3 years in the future, baselines and time-series based
approaches outperform all other methods using integrative metrics. At these time
scales environmental changes are relatively small and changes in the biota may
lag behind changes in the environment. However, as the timescale of the forecast
increases to a decade the JSDMs and environmental richness models get closer to
the accuracy of the time-series methods and even surpass some of those
methods. As the forecast distance increases the amount of environmental change
is expected to increase and the system will have more time to respond,
potentially leading to shifts towards improved relative performance of the
models incorporating environmental and ecological information. If this shift
with forecast distance in the relative importance of different processes and
models continues, this could lead to JSDMs and environmental richness models
outperforming time-series based approaches at sufficiently long time
scales. These results suggest that assessment of how forecast performance
changes with time lag is not only important for determining how far in the
future to forecast, but also for determining models and processes are most
relevant for making forecasts at different time scales.

>>>>

Forecasting is an emerging imperative in ecology and as such the field needs to
develop and follow good practices for conducting and evaluating ecological
forecasts. We have laid out a number of these practices (Box 1) and attempted to
implement them in a single study that builds and evaluates forecasts of
biodiversity in the form of species richness. The results of this effort are
both promising and humbling. On an absolute basis, when comparing forecasts
across sites, many different approaches to forecasting produce reasonable
forecasts. If a site is predicted to have a high number of species in the
future, relative to other sites, it generally does. However, the forecasts
perform relatively poorly at predicting variation in richness over time for any
given site. If a site is predicted to have a higher richness next year than it
did last year there is little confidence that this will be the case. This
results in simple site-level time-series, which are best predicted by simple
baselines like the long-term average and the "naive" model.

While stacked species distribution models are the most commonly used approach
for forecasting the future state of biodiversity [[cite?]], this approach
provided the worst overall forecasts of all methods evaluated. In fact, the most
common version of the stacking approach (based on rounding species' occurrence
probabilities up or down before adding them up) would have led to even worse
results than what we presented [Pellissier et al. 2013's "probabilistic
approach" paper, Calabese et al. 2014]. In general, SDM-based forecasts are
rarely evaluated, particularly in terms of repeated observations at the same
site, and especially with long time horizons. Since site-level richness is
relatively stable, the model evaluations that show high accuracy in predicting
differences among sites may not indicate much about the model's ability to
predict changes over time. This issue has also recently been recently
highlighted as an issue for using SDMs to predict the future locations of
individual species [plos one 2012 paper]. This result is particularly sobering
because this approach forms the foundation for major claims regarding the
predicted loss of biodiversity to climate change [2004 science/nature papers].
The poor performance of SDM-based forecasts highlights the crucial importance of
comparing multiple modeling approaches when conducting ecological forecasts, and
in particular the value of comparing results to simple baselines to avoid
over-interpreting the accuracy of ecological forecasts. Disciplines that have
more mature forecasting cultures often do this by reporting "forecast skill",
which is the improvement in the forecast relative to a simple baseline. We
recommend adopting this approach in future ecological forecasting research. It
also highlights the need for using hindcasting and time-series data for
assessing the effectiveness of forecasting methods, since this allows
comparisons to baselines and assessment of the predicted dynamics at each site.

The most common alternative to summed SDMs is modeling richness directly using
environmental variables and using the resulting models to make forecasts []. Two
comparisons of the the SDM and direct richness modeling approaches reported that
the methods yielded equivalent results for forecasting diversity (some other
comparisons that found poor SDM performance were [[tainted]] by the rounding
error alluded to above and are not considered here). While our results generally
support rough equivalence between stacked SDMs and direct richness modeling for
point estimates, they also show that modeling richness directly
results in much better estimation of the uncertainty of the forecast and
therefore is a better overall approach to forecasting richness. A similar result
is seen when comparing joint species distribution models (JSDMs) to stacked
single species distribution models. For point estimates the joint distribution
models are roughly equivalent to stacked SDMs, but the JSDMs provide much better
estimates of uncertainty. In fact, JSDMs and modeling richness directly provide
some of the best estimates of uncertainty across all modeling approaches. This
highlights the importance evaluating models based on their uncertainty as well
as their mean estimates. Not doing so is particularly problematic in this case
because the summed species distribution models are overly confident meaning that
using them would lead to believing that richness would be restricted to a much
narrower range than would actually occur.

For error metrics that evaluate both models' point estimates and their
uncertainty, the joint SDM and the nonlinear richness model began to outperform
the time-series baselines at the longest time-scales assessed. This convergence
at longer time scales highlights the importance of considering how forecasting
method performance changes with the distance into the future being predicted.
The value of this approach, and associated forecast horizons, has been raised
for knowing how far into the future a model can be effectively used for
forecasting [@petchy2015], but our results suggest a broader value to this
approach for considering the potential importance of different models and
processes for making forecasts at different scales. [[Should make sure Petchy et
al. don't make this point]] Our results show that for forecasts 1-3 years in the
future, baselines and time-series based approaches outperform all other methods
using integrative metrics. At these time scales, environmental changes are
relatively small and changes in the biota may lag behind changes in the
environment. However, as the timescale of the forecast increases to a decade the
JSDMs and environmental richness models get closer to the accuracy of the
time-series methods and even surpass some of those methods. As the forecast
distance increases the amount of environmental change is expected to increase
and the system will have more time to respond, potentially leading to shifts
towards improved relative performance of the models incorporating environmental
and ecological information. If this shift with forecast distance in the relative
importance of different processes and models continues, this could lead to JSDMs
and environmental richness models outperforming time-series based approaches at
sufficiently long time scales. These results suggest that assessment of how
forecast performance changes with time lag is not only important for determining
how far in the future to forecast, but also for determining models and processes
are most relevant for making forecasts at different time scales.
